---
title: "Some ancillary functions in package 'ipv'"
author: "F. Tusell"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(cache=TRUE, fig.width=7, fig.height=5.5)
options(width=120)
```

## Data


```{r}
library(ipv)
library(rgdal)
library(rgeos)
library(tidyverse, verbose=FALSE)
data("venta.dep")
data("alquiler.dep")
```

(*Beware*: if you are using a publicly available version of this package, such data may have been obscured, distorted and shuffled to preserve confidentiality. You will not be able to reproduce the results in this vignette, although the data provided will suffice to test
the use of functions.)

## Ancillary functions

The package contains some ancillary functions which are not meant to be called directly by the user(`pols` and
`CompFechas`) and one (`cloromap`) which affords easy representation of a variable onto the map of the CAPV (Autonomous Community of the Basque Country) or parts of it.


### cloromap

Tomemos las observaciones de inmuebles a la venta y contabilicemos el número de 
ellas que existen por cada área funcional (codificada en la variable `AF`):
Let+s take the observations corresponding to houses for sale and count the number
in each "functional are"  (coded in variable `AF`):

```{r}
Obs <- venta.dep %>%
  group_by(AF) %>%  
  summarise( Obs= n() ) %>%
  as.data.frame()
```

The  `Obs` data frame has the following structure:

```{r}
head(Obs)
```

In order to represent these numbers on a map, we need an spatial dataframe
 (spframe) with polygons for the spatial entities concerned (here, functional areas, as
 defined in `AF`). We can obtain these polygons as follows:

```{r}
dsn <- system.file("extdata/CB_AREAS_FUNCIONALES_5000_ETRS89.shp", 
                   package = "ipv")[1]
af  <- readOGR(dsn)
```

We have to check that the number and correspondence of areas in `Obs` and in `af` is consistent:

```{r}
af@data
unique(Obs$AF)
cbind(levels(af@data$A_FUNC_CAS), 
      levels(sort(unique(Obs$AF))))
```

There is one-to-one correspondence. We rename the levels in  `venta.dep`
with the names in `af`.

```{r}
levels(Obs$AF) <- levels(af@data$A_FUNC_CAS)
```

Now we can invoke:


```{r}
cloromap(sp=af, var.sp="A_FUNC_CAS", df=Obs, var.df="AF",
         var="Obs", legend.title="Obs.\ndisponibles",
         graf.title="Obs. por área funcional",
         midpoint=36000)
```

Lo que es crucial para que lo anterior funcione es la correspondencia 1-1 entre
los niveles de la variable `var.sp` de la dataframe espacial `sp`  y la variable
`var.df` de la dataframe ordinaria `df`. El argumento `midpoint` da el valor "neutro" (codificado con blanco en la escala cromática divergente empleada), que habitualmente querremos hacer coincidir con la mediana de los valores, pero podemos fijar a nuestro antojo. 

### Fusion

Quite often the names identifying locations are not homogeneous across sources, which
makes it difficult to merge information from different files. The function `Fusion` tries to simplity life a bit in such situation. It can be used in both ways, with one or two dataframes.

Whe used with one dataframe (which has to be 'X', de first argument), it scans the column
named in `varX`, performs a match with the column `locX` in the translation dataframe `canon`, and overwrites `varX` with the canonical names in `canon` (or fills an NA if a match is not possible).

This sort of invocation with a single dataframe is useful to aggregate. Assume a number of areas, some of which make part of the CAPV (Basque Autonomous Community) while others do not.
If we want to aggregate CAPV and Non-CAPV areas, we could construct a translation dataframe
`canon` as follows:

```{r}
canon <- data.frame(X=c("Bilbao","San Sebastian","Vitoria","Teruel","Soria"),
                    Y=c("Bilbao","Donostia","Vitoria/Gazteiz", "Teruel","Soria"),
                    Canon=c("CAPV","CAPV","CAPV","NoCAPV","NoCAPV"))
canon
```

Now considera a dataframe such as:

```{r}
X     <- data.frame(a=c("Bilbao","San Sebastian","Vitoria","Teruel"),
                    b=c(100,70,35,12))
X
```

In order to reclassify observations in CAPV and Non-CAPV, we can invoke:

```{r}
Fusion(X, varX="a", locX=1, locCanon=3, canon=canon)
```


We may also invoke `Fusion` with two data frames, `X` and `Y`, each of them containing, among others, a variable "naming" records: these could be, for instance, names of municipalities, of streets, of districts, or neighbourhoods.
For instance, besides `X` above we may have:

```{r}
Y     <- data.frame(e=c("Bilbao","Donostia","Vitoria/Gazteiz","Teruel","Soria"),
                    g=c("F","G","H","J","K"))
Y
```

Here the names of the communities (in column `e`) do not match those in column `a` of `X` above, although 
we are referring to the same entities (e.g., Donostia is the same as San Sebastian). Merging directly the
two dataframes `X` and `Y` would fail, but `Fusion` may be used:

```{r}
Fusion(X,Y,varX="a",varY="e",locX=1, locY=2, locCanon=2, canon=canon)
```

Above we are saying that we want to merge `X` and `Y` using the columns `a` and `e`, **after replacing their values by the canonical names in column `locCanon` of `canon`.** What this in effect does is to take the names in the dataframe
`Y` as canonical. If `locCanon` is not stated, the canonical names would be taken from the column namec `Canon` in dataframe `canon`; this would not make much sense in this case:

```{r}
Fusion(X,Y,varX="a",varY="e",locX=1, locY=2, canon=canon)
```

A translation dataframe `canon` can contain any number of columns, affording flexibility in dealing with data from different sources, in different languages, character codes, etc.
