---
title: "Nomenclature matrices"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, cache=FALSE, fig.width=7, fig.height=5.5)
options(width=120)
```


## Purpose

This vignette documents the construction of the "canonical matrices", containing the
correspondence between geographical names from differente sources. Along with funcion
`Fusion`, these matrices allow matching records referring to the same unit coming from
different sources.

### Standardized names of provinces

```{r}
canonPROVINCIA <-
  data.frame(X    = c("Alava","Guipuzcoa","Vizcaya","Álava","Guipúzcoa","Bizkaia"),
	     Canon= c("Álava","Guipúzcoa","Vizcaya","Álava","Guipúzcoa","Vizcaya"))
```


### Standardized names of municipalities

The matrix is defined as follows:

```{r}
canonMUNICIPIO <-  matrix(
  c("Abadino",                              "Abadiño",
    "Abadiño",                              "Abadiño",
    "Abaltzisketa",                         "Abaltzisketa",
    "Abanto y Ciervana/Abanto Zierbena",    "Abanto y Ciérvana-Abanto Zierbena",
    "Abanto y Ciérvana/Abanto Zierbena",    "Abanto y Ciérvana-Abanto Zierbena",
    "Abanto y Ciérvana-Abanto Zierbena",    "Abanto y Ciérvana-Abanto Zierbena",
    "Aduna",                                "Aduna",
    "Aia",                                  "Aia",
    "Aizarnazabal",                         "Aizarnazabal",
    "Ajangiz",                              "Ajangiz",
    "Albiztur",                             "Albiztur",
    "Alegia",                               "Alegia",
    "Alegria-Dulantzi",                     "Alegría-Dulantzi",
    "Alegría-Dulantzi",                     "Alegría-Dulantzi",
    "Alkiza",                               "Alkiza",
    "Alonsotegui",                          "Alonsotegi",
    "Alonsotegi",                           "Alonsotegi",
    "Altzaga",                              "Altzaga",
    "Altzo",                                "Altzo",
    "Altzola",                              "Elgoibar",
    "Amezketa",                             "Amezketa",
    "Amorebieta-Echano",                    "Amorebieta-Etxano",
    "Amorebieta-Etxano",                    "Amorebieta-Etxano",
    "Amoroto",                              "Amoroto",
    "Amurrio",                              "Amurrio",
    "Andoain",                              "Andoain",
    "Anoeta",                               "Anoeta",
    "Antzuola",                             "Antzuola",
    "Aracaldo",                             "Arakaldo",
    "Arakaldo",                             "Arakaldo",
    "Arama",                                "Arama",
    "Aramaio",                              "Aramaio",
    "Arantzazu",                            "Arantzazu",
    "Arcentales",                           "Artzentales",
    "Artzentales",                          "Artzentales",
    "Areatza",                              "Areatza",
    "Aretxabaleta",                         "Aretxabaleta",
    "Arminon",                              "Armiñón",
    "Armiñon",                              "Armiñón",
    "Armiñón",                              "Armiñón",
    "Arraia-Maeztu",                        "Arraia-Maeztu",
    "Arrankudiaga",                         "Arrankudiaga",
    "Arrasate o Mondragon",                 "Arrasate / Mondragón",
    "Arrasate / Mondragon",                 "Arrasate / Mondragón",
    "Arrasate/Mondragon",                   "Arrasate / Mondragón",
    "Arrasate/Mondragón",                   "Arrasate / Mondragón",
    "Arratzu",                              "Arratzu",
    "Arrazua-Ubarrundia",                   "Arratzua-Ubarrundia",
    "Arratzua-Ubarrundia",                  "Arratzua-Ubarrundia",
    "Arrieta",                              "Arrieta",
    "Arrigorriaga",                         "Arrigorriaga",
    "Artea",                                "Artea",
    "Artziniega",                           "Artziniega",
    "Asparrena",                            "Asparrena",
    "Asteasu",                              "Asteasu",
    "Astigarraga",                          "Astigarraga",
    "Astrabudua",                           "Erandio",
    "Ataun",                                "Ataun",
    "Ataún",                                "Ataun",
    "Atxondo",                              "Atxondo",
    "Aulesti",                              "Aulesti",
    "Ayala",                                "Ayala / Aiara",
    "Ayala / Aiara",                        "Ayala / Aiara",
    "Ayala/Aiara",                          "Ayala / Aiara",
    "Azkoitia",                             "Azkoitia",
    "Azpeitia",                             "Azpeitia",
    "Bakio",                                "Bakio",
    "Baliarrain",                           "Baliarrain",
    "Balmaseda",                            "Balmaseda",
    "Baños de Ebro",                        "Baños de Ebro / Mañueta",
    "Banos de Ebro",                        "Baños de Ebro / Mañueta",
    "Baños de Ebro / Mañueta",              "Baños de Ebro / Mañueta",
    "Barakaldo",                            "Barakaldo",
    "Barrika",                              "Barrika",
    "Barrundia",                            "Barrundia",
    "Basauri",                              "Basauri",
    "Beasain",                              "Beasain",
    "Bedia",                                "Bedia",
    "Beizama",                              "Beizama",
    "Belauntza",                            "Belauntza",
    "Berango",                              "Berango",
    "Berantevilla",                         "Berantevilla",
    "Berastegi",                            "Berastegi",
    "Bergara",                              "Bergara",
    "Bermeo",                               "Bermeo",
    "Bernedo",                              "Bernedo",
    "Berriatua",                            "Berriatua",
    "Berriz",                               "Berriz",
    "Berrobi",                              "Berrobi",
    "Bidegoyan",                            "Bidania-Goiatz",
    "Bidania-Goiatz",                       "Bidania-Goiatz",
    "Bilbao",                               "Bilbao",
    "Busturia",                             "Busturia",
    "Campezo",                              "Campezo / Kanpezu",
    "Campezo / Kanpezu",                    "Campezo / Kanpezu",
    "Campezo/Kanpezu",                      "Campezo / Kanpezu",
    "Carranza",                             "Karrantza Harana / Valle de Carranza",
    "Karrantza Harana / Valle de Carranza", "Karrantza Harana / Valle de Carranza",
    "Karrantza Harana/Valle de Carranza",   "Karrantza Harana / Valle de Carranza",
    "Karrantza Harana Valle de Carranza",   "Karrantza Harana / Valle de Carranza",
    "Cripan",                               "Kripan",
    "Cripán",                               "Kripan",
    "Kripán",                               "Kripan",
    "Cuartango",                            "Kuartango",
    "Kuartango",                            "Kuartango",
    "Deba",                                 "Deba",
    "Derio",                                "Derio",
    "Dima",                                 "Dima",
    "Donostia-San Sebastian",               "Donostia / San Sebastián",
    "Donostia-San Sebastián",               "Donostia / San Sebastián",
    "Donostia / San Sebastián",             "Donostia / San Sebastián",
    "Durango",                              "Durango",
    "Ea",                                   "Ea",
    "Eibar",                                "Eibar",
    "Elantxobe",                            "Elantxobe",
    "Elburgo",                              "Elburgo / Burgelu",
    "Elburgo/Burgelu",                      "Elburgo / Burgelu",
    "Elciego",                              "Elciego",
    "Elduain",                              "Elduain",
    "Elgeta",                               "Elgeta",
    "Elgoibar",                             "Elgoibar",
    "Elorrio",                              "Elorrio",
    "Elvillar",                             "Elvillar / Bilar",
    "Elvillar/Bilar",                       "Elvillar / Bilar",
    "Erandio",                              "Erandio",
    "Ereno",                                "Ereño",
    "Ereño",                                "Ereño",
    "Ermua",                                "Ermua",
    "Errezil",                              "Errezil",
    "Errigoiti",                            "Errigoiti",
    "Eskoriatza",                           "Eskoriatza",
    "Etxebarri, Anteiglesia de San esteban", "Etxebarri",
    "Etxebarri",                            "Etxebarri",
    "Etxebarria",                           "Etxebarria",
    "Ezkio-Itsaso",                         "Ezkio-Itsaso",
    "Forua",                                "Forua",
    "Fruiz",                                "Fruiz",
    "Gabiria",                              "Gabiria",
    "Gainza",                               "Gaintza",
    "Gaintza",                              "Gaintza",
    "Galdakao",                             "Galdakao",
    "Galdames",                             "Galdames",
    "Gallarta",                             "Abanto y Ciérvana-Abanto Zierbena",
    "Gamiz-Fika",                           "Gamiz-Fika",
    "Garay",                                "Garai",
    "Gatica",                               "Gatika",
    "Gatika",                               "Gatika",
    "Gauteguiz de Arteaga",                 "Gautegiz Arteaga",
    "Gauteguiz Arteaga",                    "Gautegiz Arteaga",
    "Gautegiz Arteaga",                     "Gautegiz Arteaga",
    "Gaztelu",                              "Gaztelu",
    "Gernika-Lumo",                         "Gernika-Lumo",
    "Getaria",                              "Getaria",
    "Getxo",                                "Getxo",
    "Gordexola",                            "Gordexola",
    "Gorliz",                               "Gorliz",
    "Guenes",                               "Güeñes",
    "Güeñes",                               "Güeñes",
    "Gueñes",                               "Güeñes",
    "Guizaburuaga",                         "Gizaburuaga",
    "Gizaburuaga",                          "Gizaburuaga",
    "Hernani",                              "Hernani",
    "Hernialde",                            "Hernialde",
    "Hondarribia",                          "Hondarribia",
    "Ibarra",                               "Ibarra",
    "Ibarranguelua",                        "Ibarrangelu",
    "Ibarranguelu",                         "Ibarrangelu",
    "Ibarrangelu",                          "Ibarrangelu",
    "Idiazabal",                            "Idiazabal",
    "Igorre",                               "Igorre",
    "Ikaztegieta",                          "Ikaztegieta",
    "Irun",                                 "Irun",
    "Iruna de Oca",                         "Iruña Oka / Iruña de Oca",
    "Iruña de Oca",                         "Iruña Oka / Iruña de Oca",
    "Iruña Oka / Iruña de Oca",             "Iruña Oka / Iruña de Oca",
    "Iruña Oka/Iruña de Oca",               "Iruña Oka / Iruña de Oca",
    "Irura",                                "Irura",
    "Iruraiz-Gauna",                        "Iruraiz-Gauna",
    "Ispaster",                             "Ispaster",
    "Itsasondo",                            "Itsasondo",
    "Iurreta",                              "Iurreta",
    "Izurza",                               "Izurtza",
    "Kortezubi",                            "Kortezubi",
    "Labastida",                            "Labastida / Bastida",
    "Labastida/Bastida",                    "Labastida / Bastida",
    "Lagran",                               "Lagrán",
    "Lagrán",                               "Lagrán",
    "Laguardia",                            "Laguardia",
    "Lanciego",                             "Lanciego / Lantziego",
    "Lanciego/Lantziego",                   "Lanciego / Lantziego",
    "Lanestosa",                            "Lanestosa",
    "Lantaron",                             "Lantarón",
    "Lantarón",                             "Lantarón",
    "Lapuebla de Labarca",                  "Lapuebla de Labarca",
    "Larrabetzu",                           "Larrabetzu",
    "Larraul",                              "Larraul",
    "Lasarte-Oria",                         "Lasarte-Oria",
    "Laukiz",                               "Laukiz",
    "Lazkao",                               "Lazkao",
    "Leaburu-Gaztelu",                      "Leaburu",
    "Leaburu"        ,                      "Leaburu",
    "Legazpia",                             "Legazpi",
    "Legazpi" ,                             "Legazpi",
    "Legorreta",                            "Legorreta",
    "Legutiano",                            "Legutio",
    "Legutio",                              "Legutio",
    "Leintz-Gatzaga",                       "Leintz-Gatzaga",
    "Leioa",                                "Leioa",
    "Lekeitio",                             "Lekeitio",
    "Lemoa",                                "Lemoa",
    "Lemoiz",                               "Lemoiz",
    "Leza",                                 "Leza",
    "Lezama",                               "Lezama",
    "Lezo",                                 "Lezo",
    "Lizartza",                             "Lizartza",
    "Llodio",                               "Laudio / Llodio",
    "Laudio/Llodio",                        "Laudio / Llodio",
    "Loiu",                                 "Loiu",
    "Mallabia",                             "Mallabia",
    "Manaria",                              "Mañaria",
    "Mañaria",                              "Mañaria",
    "Markina-Xemein",                       "Markina-Xemein",
    "Maruri",                               "Maruri-Jatabe",
    "Maruri-Jatabe",                        "Maruri-Jatabe",
    "Menaka",                               "Meñaka",
    "Mendaro",                              "Mendaro",
    "Mendata",                              "Mendata",
    "Mendexa",                              "Mendexa",
    "Menaka",                               "Meñaka",
    "Meñaka",                               "Meñaka",
    "Moreda de alava",                      "Moreda de Álava / Moreda Araba",
    "Moreda de Álava",                      "Moreda de Álava / Moreda Araba",
    "Moreda de Álava/Moreda Araba",         "Moreda de Álava / Moreda Araba",
    "Morga",                                "Morga",
    "Mundaka",                              "Mundaka",
    "Mungia",                               "Mungia",
    "Munitibar-Arbatzegi Gerrikaitz",       "Munitibar-Arbatzegi Gerrikaitz",
    "Munitibar-Arbatzegi Gerrikaitz-",      "Munitibar-Arbatzegi Gerrikaitz",
    "Murueta",                              "Murueta",
    "Muskiz",                               "Muskiz",
    "Mutiloa",                              "Mutiloa",
    "Mutriku",                              "Mutriku",
    "Muxika",                               "Muxika",
    "Nabarniz",                             "Nabarniz",
    "Navaridas",                            "Navaridas",
    "Oiartzun",                             "Oiartzun",
    "Oion",                                 "Oyón-Oion",
    "Oyón-Oion",                            "Oyón-Oion",
    "Oión",                                 "Oyón-Oion",
    "Okondo",                               "Okondo",
    "Olaberria",                            "Olaberria",
    "Onati",                                "Oñati",
    "Ondarroa",                             "Ondarroa",
    "Oñati",                                "Oñati",
    "Ordizia",                              "Ordizia",
    "Orduna",                               "Urduña / Orduña",
    "Orduña",                               "Urduña / Orduña",
    "Urduña/Orduña",                        "Urduña / Orduña",
    "Orendain",                             "Orendain",
    "Orexa",                                "Orexa",
    "Orio",                                 "Orio",
    "Ormaiztegi",                           "Ormaiztegi",
    "Orozko",                               "Orozko",
    "Ortuella",                             "Ortuella",
    "Otxandio",                             "Otxandio",
    "Pasaia",                               "Pasaia",
    "Penacerrada",                          "Peñacerrada-Urizaharra",
    "Peñacerrada",                          "Peñacerrada-Urizaharra",
    "Plentzia",                             "Plentzia",
    "Portugalete",                          "Portugalete",
    "Errenteria",                           "Errenteria",
    "Renteria / Errenteria",                "Errenteria",
    "Ribera Alta",                          "Erriberagoitia / Ribera Alta",
    "Erribera Goitia / Ribera Alta",        "Erriberagoitia / Ribera Alta",
    "Erriberagoitia/Ribera Alta",           "Erriberagoitia / Ribera Alta",
    "Ribera Baja",                          "Ribera Baja / Erribera Beitia",
    "Ribera Baja/Erribera Beitia",          "Ribera Baja / Erribera Beitia",
    "Salinas de Anana",                     "Añana",
    "Salinas de Añana",                     "Añana",
    "Añana",                                "Añana",
    "Salvatierra",                          "Agurain / Salvatierra",
    "Agurain / Salvatierra",                "Agurain / Salvatierra",
    "Agurain/Salvatierra",                  "Agurain / Salvatierra",
    "Samaniego",                            "Samaniego",
    "San Millan",                           "San Millán / Donemiliaga",
    "San Millán",                           "San Millán / Donemiliaga",
    "San Millán/Donemiliaga",               "San Millán / Donemiliaga",
    "Santurtzi",                            "Santurtzi",
    "Segura",                               "Segura",
    "Sestao",                               "Sestao",
    "Sondika",                              "Sondika",
    "Sopelana",                             "Sopela",
    "Sopela",                               "Sopela",
    "Sopuerta",                             "Sopuerta",
    "Soraluze/Placencia de las Armas",      "Soraluze-Placencia de las Armas",
    "Soraluze-Placencia de las Armas",      "Soraluze-Placencia de las Armas",
    "Sukarrieta",                           "Sukarrieta",
    "Tolosa",                               "Tolosa",
    "Trucios",                              "Turtzioz-Trucíos",
    "Trucios-Turtzioz",                     "Turtzioz-Trucíos",
    "Ubidea",                               "Ubide",
    "Ubide",                                "Ubide",
    "Ugao-Miraballes",                      "Ugao-Miraballes",
    "Urcabustaiz",                          "Urkabustaiz",
    "Urkabustaiz",                          "Urkabustaiz",
    "Urduliz",                              "Urduliz",
    "Urnieta",                              "Urnieta",
    "Urretxu",                              "Urretxu",
    "Usurbil",                              "Usurbil",
    "Valdegovia",                           "Valdegovía / Gaubea",
    "Valdegovia/Gaubea",                    "Valdegovía / Gaubea",
    "Valdegovía/Gaubea",                    "Valdegovía / Gaubea",
    "Valle de Arana",                       "Arama",
    "Valle de Trapaga-Trapagaran",          "Valle de Trápaga-Trapagaran",
    "Valle de Trápaga-Trapagaran",          "Valle de Trápaga-Trapagaran",
    "Villabona",                            "Villabona",
    "Villabuena de alava",                  "Villabuena de Álava / Eskuernaga",
    "Villabuena de Álava",                  "Villabuena de Álava / Eskuernaga",
    "Villabuena de Álava/Eskuernaga",       "Villabuena de Álava / Eskuernaga",
    "Vitoria-Gasteiz",                      "Vitoria-Gasteiz",
    "Yecora",                               "Yécora / Iekora",
    "Yécora",                               "Yécora / Iekora",
    "Yécora/Iekora",                        "Yécora / Iekora",
    "Zaldibar",                             "Zaldibar",
    "Zaldibia",                             "Zaldibia",
    "Zalduondo",                            "Zalduondo",
    "Zalla",                                "Zalla",
    "Zambrana",                             "Zambrana",
    "Zamudio",                              "Zamudio",
    "Zaratamo",                             "Zaratamo",
    "Zarautz",                              "Zarautz",
    "Zeanuri",                              "Zeanuri",
    "Zeberio",                              "Zeberio",
    "Zegama",                               "Zegama",
    "Zerain",                               "Zerain",
    "Zestoa",                               "Zestoa",
    "Zierbena",                             "Zierbena",
    "Zigoitia",                             "Zigoitia",
    "Ziortza-Bolibar",                      "Ziortza-Bolibar",
    "Zizurkil",                             "Zizurkil",
    "Zumaia",                               "Zumaia",
    "Zumarraga",                            "Zumarraga",
    "Zuya",                                 "Zuia",
    "Zuia",                                 "Zuia"),
  ncol=2, byrow=TRUE)
canonMUNICIPIO <- as.data.frame(canonMUNICIPIO)
colnames(canonMUNICIPIO) <- c("X", "Canon")
```

Here, the names in the first column are names used in idealista and other sources; the
canonical names in `Canon` are those in the cartography supplied by the Basque
Government. Even though the names do not usually match the official names and do not
follow a discernible naming convention (example: "Erriberagoitia / Ribera Alta", pero
"Ribera Baja / Erribera Beitia"), they are the names against we most often have to match
in order to construct maps.


## Saving of matrices

These matrices are saved in the `data` directory of packages `ipv`.

```{r}
save(canonPROVINCIA, file="canonPROVINCIA.rda")
save(canonMUNICIPIO, file="canonMUNICIPIO.rda")
```
